Improved Approximation Algorithms for the Maximum Carpool Matching Problme

### 2017-02-10 src/local.tex
fix
rm todos
merge dror

In this section we present a $(\frac{2c_{\max}}{c_{\max} + 1}
+ \eps)$-approximation algoithm for unweighted \carpool whose running
time is polynomial if $c_{\max} = O(1)$.

@@

\todo[inline]{fix running time paragraph}

@@

@@

@@

It remains to consider a maximal component $\calT$ such that If
$\abs{M \cap A(\calT)} > k$.  Observe that since the number of edges
in a $S \in \calS(M^*)$ is at most $\cmax$, it must be that
$\abs{V(\calT)} > \frac{k}{\cmax}$.
%
Given such a component $\calT$, we partition its vertex set into
several sets, each of size at least $\frac{k}{4\cmax}$.  We do this as
follows.  First construct any spanning tree of $\calT$ (say, by
running DFS).  It is well known that a tree of $t$ vertices can be
split into two subtrees each with at most $\floor{2t/3}+2$ vertices
(see, e.g., \cite{???}).  Moreover, if $t \geq 24$, each such subtree
contains at most $\frac{3}{4}t$.  Hence, we can decompose $\calT$
recursively into vertex sets whose sizes is between $\frac{k}{\cmax}$
and $\frac{k}{4\cmax}$, assuming that $\frac{k}{\cmax} \geq 24$.
%
Each such vertex set $\calX$ is connected and contains at most
$\frac{k}{\cmax}$ stars.  Hence $\abs{A(\calX)} \leq k$.  Due to
Lemma~\ref{lemma:r} and since $\abs{V(\calX)} \geq \frac{k}{4\cmax}$,
we have that
\[
\frac{\deg_M(V(\calX))}{\deg_{M^*}(V(\calX))} 
\geq \half + \inv{2\cmax} - \frac{2}{k}
~.
\]
Since $\deg_{M^*}(V(\calT)) = \sum_{\calX} \deg_{M^*}(V(\calX))$ and
$\deg_M(V(\calT)) = \sum_{\calX} \deg_M(V(\calX))$, it follows that
\[
\frac{\deg_M(V(\calT))}{\deg_{M^*}(V(\calT))}
\geq \half + \inv{2\cmax} - \frac{2}{k}
~,
\]
and thus $\abs{M \cap A(\calT)} \geq (\frac{\cmax + 1}{2\cmax}
- \frac{2}{k}) \abs{M^* \cap A(\calT)}$.
\end{proof}

\begin{corollary}
\label{cor:local}
There exists a $(\frac{2\cmax}{\cmax+1} +\eps)$-approximation
algorithm for \carpool, for every $\eps>0$.
\end{corollary}

@@

\todo[inline]{Add a proper theorem}

We now show that Corollary~\ref{cor:local} is tight.  Consider the
example in Figure~\ref{fig:local search tight}, the example is for the
special case when $k=11$ and $C=2$ but this example can be generalized
in a straightforward manner.  One can verify that as the graph in the
example growth, the ratio between the optimal solution and the local
search solution approaches $\frac{3}{4}$.

\begin{figure}
\begin{center}
\input{fig-local-tight}
\caption{A tight example for Theorem~\ref{theorem:local search}. 
The optimal solution is given by the red solid arcs while the local
search solution is given by the green, dashed arcs.  The solution can
be improved by the local search algorithm only if it removes all the
edges.}
\label{fig:local search tight}
\end{center}
\end{figure}

>>>>

In this section we present a $(\half + \inv{2c_{\max}}
- \eps)$-approximation algoithm for unweighted \carpool whose running
time is polynomial if $c_{\max} = O(1)$.

@@

%\todo[inline]{fix running time paragraph}

@@

Define $\deg_M(v) \eqdf \din[M](v) + \dout[M](v)$.  For a subset
$U \subseteq V$ of vertices define $\deg_M(U) \eqdf \sum_{v \in
U} \deg_M(u)$.

@@

\dror{fix proof of next lemma}

@@

It remains to consider a maximal component $\calT$ such that If
$\abs{M \cap A(\calT)} > k$.  Observe that since the number of edges
in a $S \in \calS(M^*)$ is at most $\cmax$, it must be that
$\abs{V(\calT)} > \frac{k}{\cmax}$.
%
Given such a component $\calT$, we partition its vertex set into
several sets, each of size at least $\frac{k}{4\cmax}$.  We do this as
follows.  First construct any spanning tree of $\calT$ (say, by
running DFS).  It is well known that a tree of $t$ vertices can be
split into two subtrees each with at most $\floor{2t/3}+2$ vertices
% (see, e.g., \cite{???}).  
Moreover, if $t \geq 24$, each such subtree
contains at most $\frac{3}{4}t$.  Hence, we can decompose $\calT$
recursively into vertex sets whose sizes is between $\frac{k}{\cmax}$
and $\frac{k}{4\cmax}$, assuming that $\frac{k}{\cmax} \geq 24$.
%
Each such vertex set $\calX$ is connected and contains at most
$\frac{k}{\cmax}$ stars.  Hence $\abs{A(\calX)} \leq k$.  Due to
Lemma~\ref{lemma:r} and since $\abs{V(\calX)} \geq \frac{k}{4\cmax}$,
we have that
\[
\frac{\deg_M(V(\calX))}{\deg_{M^*}(V(\calX))} 
\geq \half + \inv{2\cmax} - \frac{2}{k}
~.
\]
Since $\deg_{M^*}(V(\calT)) = \sum_{\calX} \deg_{M^*}(V(\calX))$ and
$\deg_M(V(\calT)) = \sum_{\calX} \deg_M(V(\calX))$, it follows that
\[
\frac{\deg_M(V(\calT))}{\deg_{M^*}(V(\calT))}
\geq \half + \inv{2\cmax} - \frac{2}{k}
~,
\]
and thus $\abs{M \cap A(\calT)} \geq (\frac{\cmax + 1}{2\cmax}
- \frac{2}{k}) \abs{M^* \cap A(\calT)}$.
\end{proof}

\begin{corollary}
\label{cor:local}
There exists a $(\half + \inv{2\cmax} - \eps)$-approximation algorithm
for \carpool, for every $\eps>0$.
\end{corollary}

@@

\dror{make sure that example is aligned with the conditions of the
previous lemma}

We now show that Corollary~\ref{cor:local} is tight.  Consider the
example in Figure~\ref{fig:local search tight}, the example is for the
special case when $k = 11$ and $\cmax = 2$ but this example can be
generalized in a straightforward manner.  One can verify that as the
graph in the example growth, the ratio between the optimal solution
and the local search solution approaches $\frac{3}{4}$.

\begin{figure}
\begin{center}
\input{fig-local-tight}
\caption{A tight example for Corollary~\ref{cor:local}. 
The optimal solution is given by the red solid arcs while the local
search solution is given by the green, dashed arcs.  The solution can
be improved by the local search algorithm only if it removes all the
edges.}
\label{fig:local search tight}
\end{center}
\end{figure}

\dror{Mention [7]}

### 2017-02-10 src/sub.tex
no errors
merge dror

In this section we show that the \carpool problem can be formulated as
unconstrained submodular maximization problem, and thus it has a
randomized $\frac{1}{2}$-approximation algorithm based on~\cite{BFNS15}.

Given a \carpool instance $(G = (V,A), c, w)$, consider a subset
$S \subseteq V$.  Let $M(S)$ be the maximum weight carpool matching
that satisfies $D_{M(S)} \subseteq S \subseteq V \setminus P_{M(S)}$,
namely $M(S)$ is the best carpool matching whose drivers belong to $S$
and whose passengers belong to $V \setminus S$.  In other words,
$M(S)$ is the maximum weight carpool matching that is a subset of
$A \cap (V \setminus S) \times S$.
%
Given $S$, the carpool matching $M(S)$ can be computed in polynomial
time using a maximum b-matching in the bipartite graph 
$B = (S \cup V \setminus S, A \cap (V \setminus S) \times S)$ or can be reduced
to a max-cost flow problem as shown in~\cite{kutiel2016}.

Consider the function $\barw: 2^V \to \R$, where $\barw(S) \eqdf
w(M(S))$.  In the next lemma we prove that $\bar{w}$ is
a \emph{submodular set function}.  Recall that a function $f$ is
submodular if $f(S) + f(T) \geq f(S \cup T) + f(S \cap T)$ for every
two sets $S$ and $T$ in the domain of $f$.

@@

Buchbinder et al.~\cite{BFNS15,buchbinder2016deterministic} presented a
general randomized $2$-approximation algorithm for unconstrained submodular maximization.

\begin{theorem}
There exists a polynomial time $2$-approximation algorithm for \carpool.
\end{theorem}

\begin{figure}
\centering
\input{fig-submodular}
\caption[]{
\label{fig:defs}
We show that the arcs from \subref{sub:cup} and \subref{sub:cap} can be
redistributed over \subref{sub:s} and \subref{sub:s} 
%\todo{write explanation}
}
\end{figure}

>>>>

In this section we show that the \carpool problem can be formulated as
unconstrained submodular maximization problem, and thus it has a
$\frac{1}{2}$-approximation algorithm based on Buchbinder et
al.~\cite{BFNS15,buchbinder2016deterministic}.

Given a \carpool instance $(G = (V,A), c, w)$, consider a subset
$S \subseteq V$.  Let $M(S)$ be the maximum weight carpool matching
that satisfies $D_{M(S)} \subseteq S \subseteq V \setminus P_{M(S)}$,
namely $M(S)$ is the best carpool matching whose drivers belong to $S$
and whose passengers belong to $V \setminus S$.  In other words,
$M(S)$ is the maximum weight carpool matching that is a subset of
$A \cap (V \setminus S) \times S$.
%
Given $S$, the carpool matching $M(S)$ can be computed in polynomial
time by computing a maximum $b$-matching in the bipartite graph $B =
(S \cup V \setminus S, A \cap (V \setminus S) \times S)$ which can be
done using an algorithm for \textsc{Minimum Cost Flow} as shown
in~\cite{kutiel2016}.

Consider the function $\barw: 2^V \to \R$, where $\barw(S) \eqdf
w(M(S)) = \sum_{e \in M(S)} w(e)$.  Observe that $\barw(\emptyset)
= \barw(V) = 0$, and that $\barw$ is not monotone.
%
In the next lemma we prove that $\bar{w}$ is a \emph{submodular set
function}.  Recall that a function $f$ is submodular if $f(S) +
f(T) \geq f(S \cup T) + f(S \cap T)$ for every two sets $S$ and $T$ in
the domain of $f$.

@@

Buchbinder et al.~\cite{BFNS15,buchbinder2016deterministic} presented
a general randomized $\half$-approximation algorithm for unconstrained
submodular maximization.

\begin{theorem}
There exists a polynomial time $\half$-approximation algorithm
for \carpool.
\end{theorem}

\begin{figure}
\centering
\input{fig-submodular}
\caption[]{
We show that the arcs from \subref{sub:cup} and \subref{sub:cap} can be
redistributed over \subref{sub:s} and \subref{sub:s} 
%\todo{write explanation}
}
\end{figure}
